---
title: "Loot Boxes in Video Games: Analysis and Insights"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This R Markdown file outlines the steps taken to analyze survey data related to loot boxes in video games. The analysis focuses on trends in spending, perceptions of gambling, and support for regulation.

---

## Step 1: Load Libraries and Data
```{r libraries, warning=FALSE}
# Load required libraries
library(dplyr)
library(ggplot2)
library(tidyverse)
```

```{r load-data}
# Load the dataset
data <- read.csv("Survey_Results.csv")

# Rename columns for better readability
colnames(data) <- c(
  "Agreement", "Familiarity", "PurchasedLootBox", "EngagementFrequency",
  "YearlySpending", "NegativeFinancialImpact", "PurchaseFactors", 
  "OtherFactors", "ImpactOnExperience", "PressureToPurchase", 
  "AddictiveBehavior", "ConsideredGambling", "RegulationSupport", 
  "AgeRestrictionsAwareness", "AgeLimitSupport", "OddsDisclosure", 
  "StricterRegulation", "ObservedGamblingProblems"
)
```

---

## Step 2: Data Cleaning
```{r data-cleaning}
# Handle missing values and standardize responses
data <- data %>%
   mutate(across(where(is.character), ~ifelse(. == "" | . == " ", "No Response", .))) %>%
  mutate(across(where(is.character), ~replace_na(., "No Response")))

#Remove rows for participants who did not agree
data <- data[data$Agreement != "Do not Agree",]

```

---

## Step 3: Analyze Trends

### Spending Patterns
```{r spending-analysis}
# Scatter plot for spending vs. frequency

#Filter out people who haven't spent any money
filtered_data = data %>% 
  filter(YearlySpending != "Have not spent any money")

ggplot(filtered_data, aes(x = YearlySpending, y = EngagementFrequency)) +
  geom_jitter(color = "purple", size = 2, alpha = 0.6) +
  labs(title = "Spending vs. Engagement Frequency",
       x = "Yearly Spending (\u00A3)",
       y = "Engagement Frequency") +
  theme_minimal()


```

### Familiarity with Loot Boxes
```{r familiarity-analysis}
# Pie chart for familiarity levels
familiarity_dist <- data %>%
  group_by(Familiarity) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = round((Count / sum(Count)) * 100, 2))

ggplot(familiarity_dist, aes(x = "", y = Percentage, fill = Familiarity)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y") +
  labs(title = "Familiarity with Loot Boxes",
       fill = "Familiarity Level") +
  geom_text(aes(label = paste0(Percentage, "%")),            # Add percentage labels
            position = position_stack(vjust = 0.5),          # Center labels in slices
            size = 4, color = "white")+ 
  theme_void()

```

### Perceptions of Gambling
```{r gambling-analysis}
# filtering and sorting the data
gambling_opinion <- data %>%
  filter(ConsideredGambling != "No Response") %>%
  group_by(ConsideredGambling) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = round((Count / sum(Count)) * 100, 2))


#Visualize perception of gambling
ggplot(gambling_opinion, aes(x = ConsideredGambling, y = Percentage, fill = ConsideredGambling)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Do Participants Consider Loot Boxes as Gambling?",
       x = "Response",
       y = "Percentage") +
  theme_minimal()
```

---

### Support for Regulation
```{r regulation-analysis}
#filter and group the data while also changing a column entry so it fits in the chart
regulation_support <- data %>%
  filter(StricterRegulation != "No Response") %>%
  group_by(StricterRegulation) %>%
  summarise(Count = n()) %>%
  mutate(Percentage = round((Count / sum(Count)) * 100, 2)) %>% 
  mutate(StricterRegulation = gsub("No, the current level of regulation is sufficient", "Regulation is already sufficient", StricterRegulation)) 


# Visualize regulation support
ggplot(regulation_support, aes(x = StricterRegulation, y = Percentage, fill = StricterRegulation)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Support for Stricter Regulation of Loot Boxes",
       x = "Regulation Level",
       y = "Percentage") +
  theme(axis.text.x = element_text(angle =50, vjust = 1, hjust=1))
```
### Familiarity vs Regulation support
```{r}
# filtering and grouping the data before visualization
heatmap_data <- data %>%
  filter(Familiarity != "No Response", RegulationSupport != "No Response") %>%
  group_by(Familiarity, RegulationSupport) %>%
  summarise(Count = n())

#visualizing filtered data
ggplot(heatmap_data, aes(x = Familiarity, y = RegulationSupport, fill = Count)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Heatmap of Familiarity vs. Regulation Support",
       x = "Familiarity Level",
       y = "Regulation Support",
       fill = "Count") +
  theme_minimal()


```

---

## Conclusion
This analysis provides insights into spending patterns, perceptions of gambling, and regulatory opinions regarding loot boxes in video games. These findings can guide stakeholders in making informed decisions about game design, monetization strategies, and policies.


